---
title: "ToothGrowth_Basic Inferential Data Analysis"
author: "Yato"
date: "March 16, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This part is going to analyze the ToothGrowth data in the R datasets package.

1. Load the ToothGrowth data, and provide a basic summary of the data.
2. Perform some basic exploratory data analyses.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
4. State your conclusions and the assumptions needed for your conclusions.

## Loading the data and exploratory data analyses

```{r, echo = FALSE, results = "hide", warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(HH)
library(gridExtra)
```
```{r}
data(ToothGrowth); attach(ToothGrowth)
```

## Basic Summary of the Data

```{r}
str(ToothGrowth)
summary(ToothGrowth)
unique(dose)
table(supp, dose)
ToothGrowth %>% group_by(supp, factor(dose)) %>% summarise(mean_len = mean(len), sd_len = sd(len))
```

## Exploratory Data Analysis
```{r, fig.height=2}
gg1<-ggplot(ToothGrowth, aes(x = factor(dose), y = len, fill = supp)) + geom_boxplot() 
gg2 <- ggplot(ToothGrowth, aes(x = supp, y = len, fill = factor(dose))) + geom_boxplot() 
grid.arrange(gg1, gg2, ncol = 2) # or interaction2wt(len ~ supp*factor(dose))
```

## Confidence Interval and Hypothesis Testing
* Examining the impact of **supp** or **dose** as a factor separately.
* Examining the impact of supp, dose and the interaction of supp and dose by including both **supp and dose** as factors.
```{r}
dose15 <- ToothGrowth %>% filter(dose %in% c("0.5", "1"))
dose12 <- ToothGrowth %>% filter(dose %in% c("1", "2"))
dose52 <- ToothGrowth %>% filter(dose %in% c("0.5", "2"))
t.test(len ~ supp)
t.test(len ~ factor(dose), data = dose15)
t.test(len ~ factor(dose), data = dose12)
t.test(len ~ factor(dose), data = dose52)
# or summary(aov(len ~ dose)) and summary(aov(len ~ supp*factor(dose)))
detach(ToothGrowth)
```

## Conclusion
* This is a balanced design(equal sample sizes in each cell of the design).
* The confidence interval is [-0.171, 7.571], which include zero. There is no significant correlation between delivery methods and tooth length.
* The confidence interval is [-11.98, -6.276] for doses 0.5 and 1.0, [-18.16, -12.83] for doses 0.5 and 2.0, and [-8.996, -3.734] for doses 1.0 and 2.0) and show that there is a significant correlation between tooth length and dose levels.
* T-test shows that two delivery methods (OJ or VC) have equal tooth length (p>0.05). Three levels of dose aren't all equally effective. Increasing the dose level can lead to increased tooth growth.
* Both main effects and the interaction between these factors are significant.



